DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-18 paper/evaluation-cachepart.tex
revision up to 4.H

Cache partitioning is a well-known technique to improve isolation in
real-time systems by giving a dedicated cache space to individaul task
or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
page-coloring based cache partitioning solution for Linux, to
investigate the effect of cache partitioning in protecting DeepPicar's
DNN based controller.

@@

Figure \ref{fig:palloc_bandwidth_exectime} shows the
result. For BwRead, when cache partitioning is enabled, DNN
performance becomes slightly better, with a improvement of  
\textasciitilde6.5\%. For BwWrite, cache partitioning actually makes
the DNN performance slightly worse that that of w/o partitioning. In
either case, the main observation is that cache partitioning is not
effective in providing timing isolation to our DNN workload.

>>>>

Cache partitioning is a well-known technique to improve isolation in
real-time systems by giving a dedicated cache space to each individaul
task or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
page-coloring based cache partitioning solution for Linux, to
investigate the effect of cache partitioning in protecting DeepPicar's
DNN based controller.

@@

Figure \ref{fig:palloc_bandwidth_exectime} shows the
result. For BwRead, when cache partitioning is enabled, DNN
performance becomes slightly better, with a improvement of  
\textasciitilde6.5\%. For BwWrite, cache partitioning actually makes
the DNN performance slightly worse that that of w/o partitioning.

In summary, we find that the DNN inferencing workload is not sensitive
to cache space and that cache partitioning is not effective in
providing timing isolation to our DNN workload.

### 2018-04-18 paper/evaluation-drampart.tex
revision up to 4.H

To determine the how sensitive the model is to the shared DRAM 
memory, we use MemGuard\cite{yun2014rtas}, a memory bandwidth reservation system that 
can limit the amount of bandwidth each core receives in a period. For 
all tests run using MemGuard, we use a period of one second.

@@

We first run a single model on one core, Core 0, while varying the 
core's bandwidth reservation size, from 500 MB/s down to 100 MB/s. 
The results can be seen in Figure \ref{fig:memguard_multicore}. When less 
memory bandwidth is available to the core the model's performance 
noticably decreases.
%% , and it performs the best when no memory bandwidth 
%% throttling is implemented. 
Compared to a reservation size of 100 
MB/s, disabling memory bandwidth throttling results in 
\textasciitilde3x performance improvement. If the model was memory 
insensitive, then the amount of bandwidth available to it would have 
little effect on its performance. However, since that is not the 
case, we conclude that, to a notable extent, the neural network model 
is dependent on shared DRAM memory performance.

@@

We also test the performance of the model in the presence of three
memory intensive read and write benchmarks to determine the effects of 
limiting the memory bandwidth of the co-runners. For each of the 
experiments, we give the model's core, Core 0, a constant 1000 MB/s 
memory bandwidth reservation size, and then vary the amount of bandwidth 
the co-runners' cores receive. For read co-runners, we give them an initial
reservation size of 500 MB/s, and decrease that to a minimum of 100 MB/s.
In the case of write co-runners, MemGuard accounts for L2 refills but not
write-backs, so each benchmark would effectively have twice the bandwidth
of their reservation size. Due to this, we give write co-runners an initial 
reservation size of 250 MB/s and decrease it to a minimum of 50 MB/s, but
their effective bandwidths are the same as the read co-runners.
The results can be see in Figures \ref{fig:memguard_bandwidth}. 
When less memory bandwidth is given to both the read and write 
co-runners, the performance of the model improves. 
%% In the case of one co-runner, the performance of the model remains 
%% constant even as the reservation sizes are decreased, however, it 
%% still improves compared to when throttling wasn't enabled. 
In both cases, when the co-runners were given a minimal reservation size,
the performance of the model was much closer to its solo execution time.
This is especially noteworthy when write co-runners were present, as the 
model didn't suffer the same 11.6X slowdown that was seen without
bandwidth throttling enabled. 
%% When memory intensive Bandwidth benchmarks were present, model performance
%% improved, to a notable extent, through the use of memory bandwidth 
%% throttling on the co-runners. 
Since this would not be the case if the 
model was memory insensitive, we find that the model is memory 
dependent in the presence of co-runners.

Based on all of the memory bandwidth throttling experiments 
performed, it is evident that the shared DRAM memory is important for
the DNN model. When a single model was run, its 
performance decreased as less memory bandwidth was 
made available to it. Furthermore, when the bandwidths of memory 
intensive read and write co-runners were limited, the performance of 
the model improved. As such, we conclude that the model is dependent on the 
shared DRAM memory.

>>>>

In this subsection, we examine the DNN workload's memory bandwidth
sensitivity and the effect of memory bandwidth throttling in providing
isolation. For the experiments, we use MemGuard\cite{Yun2013}, which 
can limit the amount of memory bandwidth each core receives. MemGuard
operates periodically, at an 1ms interval, and uses hardware
performance counters to throttle cores if they exceed their given
bandwidth budgets within each regulation period (i.e., 1ms) by
scheduling high-priority idle kernel threads.

In the first experiment, we measure the performance of the DNN model
running on a single core, Core 0, first w/o using MemGuard and then w/
using MemGuard while varying the core's bandwidth throttling parameter
from 500 MB/s down to 100 MB/s.

@@

Figure \ref{fig:memguard_multicore} shows the result. When the DNN
model executing core is throttled at 400 MB/s or more, performance of
the model is largely the same as the non-throttled case. However, as
we decreate the assigned memory badnwdith below 300 MB/s, we start to
observe noticably decreases in the model's performance. In other
words, the DNN model is sensitive to memory bandwidth and that it
requires 400 MB/s or more bandwidth should be given to ensure ideal
performance.

In the next experiment, we again repeat the experiment in
Section~\ref{sec:eval-memhog}---i.e., co-scheduling the DNN model and
three Bandwidth (BwRead or BwWrite) instances---but this time we
throttle the cores that execute the co-runners using MemGuard by 
varying their bandwidth budgets to see ther impact in the DNN model's
performance.

Figures \ref{fig:memguard_bandwidth} shows the result.
Note that, in case of BwWrite co-runners, we have assigned half the
bandwidth budget because MemGuard currently only account L2 refills
but not write-backs, which effectively allow twice the allocated
memory bandwidth budget. Regardless, the result clearly shows that
limiting co-runners's memory bandwidth is effective in protecting the
DNN model's performance.

In summary, we find that the DNN inferencing workload is sensitive to
memory bandwidth and that memory bandwidth throttling is effective in
improving performance isolation of the DNN workload.

@@

%% In the case of one co-runner, the performance of the model remains 
%% constant even as the reservation sizes are decreased, however, it 
%% still improves compared to when throttling wasn't enabled. 
%% In both cases, when the co-runners were given a minimal reservation size,
%% the performance of the model was much closer to its solo execution time.
%% This is especially noteworthy when write co-runners were present, as the 
%% model didn't suffer the same 11.6X slowdown that was seen without
%% bandwidth throttling enabled. 
%% %% When memory intensive Bandwidth benchmarks were present, model performance
%% %% improved, to a notable extent, through the use of memory bandwidth 
%% %% throttling on the co-runners. 
%% Since this would not be the case if the 
%% model was memory insensitive, we find that the model is memory 
%% dependent in the presence of co-runners.

%% Based on all of the memory bandwidth throttling experiments 
%% performed, it is evident that the shared DRAM memory is important for
%% the DNN model. When a single model was run, its 
%% performance decreased as less memory bandwidth was 
%% made available to it. Furthermore, when the bandwidths of memory 
%% intensive read and write co-runners were limited, the performance of 
%% the model improved. As such, we conclude that the model is dependent on the 
%% shared DRAM memory.

### 2018-04-18 paper/evaluation.tex
revision up to 4.H

The Raspberry Pi 3 Model B platform used in DeepPicar equips a Broadcom
BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has a 32K private Instruction cache
and a 32K private Data cache, and all cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV GPU, although we did
not use the GPU in our evaluation, due to the lack of sofware support
\footnote{TensorFlow currently only support NVIDIA's GPUs.}.
For software, we use Ubuntu MATE 16.04, TensorFlow 1.1.
We disabled DVFS (dynamic voltage frequency scaling) and
configured the clock speed of each core statically at the maximum 1.2GHz.

@@

Given the importance of predictable timing on real-time control, such
as our DNN based control task, we wanted to understand (1) the main
source of the timing and (2) to evaluate existing isolation methods to
avoid this kind of timing interference. Specifically, shared cache
space and DRAM bandwidth are two most well-known sources of contention
in multicore. Thus, in the following, we investigate test if and how
they influence the observed timing interference and how effective are
the existing mitigation methods. To our surprise, we find that neither
the cache space nor DRAM bandwidth is the root cause of the
DeepPicar's observed worst-case execution time increase.

@@

First, we find that DeepPicar's computing platform,
the Raspberry Pi 3 Model B, offers computing capacity to
perform real-time control of the RC car at 30 Hz frequency (or
33.$\overline{\mbox{33}}$ ms  per control loop). Given the complexity of 
the DNN used, we were pleasantly surprised by this finding. 
The time breakdown shows that the DNN inferencing operation, 
performed by the Tensorflow library, dominates the execution time, 
which is expected.

@@

Third, we find that consolidating multiple DNN models---on different CPU
cores---is feasible as we find: (1) DNN performance using a single
core is not much worse than using multiple cores; (2) multiple DNN
models running simultaneously do not cause severe interference with
each other.

Fourth, we find that consolidating memory (DRAM) performance
intensive applications could jeopadize DNN performance, because DNN
performance appears to be very sensitive to memory performance; we 
observe up to 11.6X slowdown in DNN performance by co-scheduling 
synthetic memory bandwidth intensive applications on idle cores.

Fifth, we find that partitioning the shared L2 cache provides 
practically no benefits to the real-time performance of the DNN. To the 
contrary, performance is constant regardless of the amount of cache 
space used. Furthermore, we find that performance remains consistent 
when memory intensive read co-runners are introduced, even though
the model has its own isolated cache space. As such, we 
conclude that the shared L2 cache is not essential to the DNN's real-time 
performance.

Lastly, we find that memory bandwidth throttling results in significant
changes to model performance. When less bandwidth is given to the model,
we find that it performs worse. When memory intensive co-runners are
present, we find that limiting the available bandwidth of the co-runners
results in significantly improved model execution times. As such, we conclude
that the shared DRAM memory is essential to the DNN's real-time 
performance.

>>>>

The Raspberry Pi 3 Model B platform used in DeepPicar equips a Broadcom
BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has a 32K private Instruction cache
and a 32K private Data cache, and all cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV GPU, although we did
not use the GPU in our evaluation, due to the lack of sofware support
\footnote{TensorFlow currently only supports NVIDIA's GPUs.}.
For software, we use Ubuntu MATE 16.04, TensorFlow 1.1.
We disabled DVFS (dynamic voltage frequency scaling) and
configured the clock speed of each core statically at the maximum 1.2GHz.

@@

Given the importance of predictable timing on real-time control, such
as our DNN based control task, we wanted to understand (1) the main
source of the timing and (2) to evaluate existing isolation methods to
avoid this kind of timing interference. Specifically, shared cache
space and DRAM bandwidth are two most well-known sources of contention
in multicore. Thus, in the following, we investigate if and how
they influence the observed timing interference and how effective are
the existing mitigation methods. To our surprise, we find that neither
the cache space nor DRAM bandwidth is the root cause of the
DeepPicar's observed worst-case execution time increase.

@@

First, we find that DeepPicar's computing platform, the Raspberry Pi
3, offers sufficient computing capacity to perform deap-learning based
real-time control at 30 Hz frequency (or 33.3 ms per control
loop). Given the complexity of the DNN used, we were pleasantly
surprised by this finding. 
%% The time breakdown shows that the DNN inferencing operation, 
%% performed by the Tensorflow library, dominates the execution time, 
%% which is expected.

@@

Third, we find that consolidating multiple DNN models---on different CPU
cores---is feasible as we find: (1) DNN performance using a single
core is not much worse than using multiple cores; (2) multiple DNN
models running simultaneously cause moderate interference.

Fourth, we find that consolidating memory (DRAM) performance
intensive applications (especially write-heavy ones) can jeopadize
the DNN inferencing task's performance even when all tasks have
dedicated cores due to contention in shared hardware resources; we
observe up to 11.6X slowdown.

Lastly, we find that the DNN inferencing workload is not sensitive
to cache space and that cache partitioning is not effective in
providing timing isolation to our DNN workload. On the other hand, we
find that the DNN workload is sensitive to memory bandwidth and that
memory bandwidth throttling is effective in improving performance
isolation.

%% To the 
%% contrary, performance is constant regardless of the amount of cache 
%% space used. Furthermore, we find that performance remains consistent 
%% when memory intensive read co-runners are introduced, even though
%% the model has its own isolated cache space. As such, we 
%% conclude that the shared L2 cache is not essential to the DNN's real-time 
%% performance.

%% Lastly, we find that memory bandwidth throttling results in significant
%% changes to model performance. When less bandwidth is given to the model,
%% we find that it performs worse. When memory intensive co-runners are
%% present, we find that limiting the available bandwidth of the co-runners
%% results in significantly improved model execution times. As such, we conclude
%% that the shared DRAM memory is essential to the DNN's real-time 
%% performance.
